---
title: "JI Chunyan"
output:
  html_document:
    df_print: paged
  pdf_document:
    number_sections: yes
editor_options: 
  markdown: 
    wrap: 72
---

# Introduction

This project, inspired by renowned statistician Hans Rosling, examines
the change in fertility rates from 1990 to 2020 and explores variations
in rates among countries based on the UN Human Development Index (HDI).
The project can be accessed on my GitHub repository at
[**https://github.com/Chunyan94/projectADD1.git**](https://github.com/Chunyan94/projectADD1.git).

**Source data**

1.  *United Nations Demographic Indicators*: This data contains
    information on fertility rate changes for countries from 1950 to
    2100. 
2.  *United Nations Human Development Index (HDI)*: This composite index
    measures average achievement in three dimensions of human
    development: a long and healthy life, knowledge, and a decent
    standard of living.
3.  *Gross national income (GNI) per capita* : The GNI per capita is the
    dollar value of a country's final income in a year, divided by its
    population. It should be reflecting the average before tax income of
    a country's citizens.

# Import and Cleaning of Data

## 1. UN Demographic Indicators

**File name**: "1950-2100, medium (ZIP, 7.77 MB)"

**Source link**:
[**https://population.un.org/wpp/Download/Standard/CSV/**](https://population.un.org/wpp/Download/Standard/CSV/)

**Variables**

-   LocID: area codes in M49 classification

-   ISO3_code: area codes in ISO 3166 classification

-   Location: country or area names

-   Time: year

-   TPopulation1July: total population

-   TFR: Total Fertility Rate (live births per woman)

```{r, message=FALSE}
library(readr)

popu_raw <- read_csv("https://raw.githubusercontent.com/Chunyan94/projectADD1/main/data/WPP2022_Demographic_Indicators_Medium.csv",show_col_types = FALSE)

```

### Clean data

We selected six variables from the population demographic data for this
analysis.

```{r, message=FALSE}
library(tidyverse)

popu <- popu_raw %>% 
  select(ISO3_code, LocID, Location, Time, TPopulation1July, TFR, PopGrowthRate, LEx) %>% 
  rename(iso3 = ISO3_code,
         year = Time,
         country = Location,
         population = TPopulation1July,
         fertilityRate = TFR) %>%
  arrange(LocID) 

head(popu, 5)

```

## 2. Human Development Index

**File name**: HDR21-22_Composite_indices_metadata.csv

**Source link**:
<https://hdr.undp.org/data-center/documentation-and-downloads>

**Variables**

-   iso3: area codes in ISO 3166 classification

-   country: country names

-   hdicode: four human development classification

    -   "low" : less than 0.550 for human development

    -   "medium": 0.550--0.699 for human development

    -   "high" : 0.700--0.799 for human development

    -   "very high": 0.800 or greater for human development.

-   region: region classfication

-   hdi_1990 - hdi_2021: human development index from 1990 to 2021

```{r}
HDI_raw <- read_csv("https://raw.githubusercontent.com/Chunyan94/projectADD1/main/data/HDI.csv",show_col_types = FALSE)
head(HDI_raw,3)
```

**Transpose data**

```{r}

HDI <- HDI_raw %>% 
  # select country & hdi_990 to hdi_2020
  select(c(1:4,6:36)) %>% 
  # transpose hdi_1990 - hdi_2020 
  pivot_longer(cols = starts_with("hdi_"),
               names_to = "year",
               names_prefix = "hdi_",
               values_to = "hdi") %>% 
    #  deleting empty string and NA
  filter(!is.na(hdi) & !is.na(hdicode) & hdicode != "") %>% 
  mutate(year = as.integer(year),
         hdicode = factor(hdicode, levels = c('Low', 'Medium', 'High', 'Very High')),
         hdi = as.numeric(hdi))

head(HDI,5)

```

## Merge population and HDI data

```{r}
final <- merge(popu, HDI, by = c("iso3", "year")) %>%
  # change the popu$country name 
  rename(country = country.x) %>%
  # drop extra country variable
  select(- country.y) 

head(final, 5)
```

# Data visualization

## Create World Maps

I used coordinates data from the 'maps' package and plotted them using
the 'ggplot2' library's 'map_data' function to create visual
representations of the data.

```{r, warning=FALSE}

# world map data 
coordinate <- map_data("world")

Year <- 2020

world <- final %>%
  merge(coordinate, by.x = "country", by.y = "region") %>%
  arrange(group, order) %>%
  filter(year == Year)  

ggplot(data = world) +
  geom_map(
    aes(map_id = region, 
        x = long, 
        y = lat, 
        fill=PopGrowthRate), 
    map = world
  ) +
  labs(title = paste("Population growth rate in", Year)) +
  theme_void()  
```

However, the names of regions in the map data do not align with those in
the merged "final" dataset. Therefore, it is necessary to replace the
mismatched region names.

### 1. Identify unmatched country names

Some countries does , like "United States of American". I need to
explore why there is a difference. Identify corresponding names for
countries in the final data set that do not match the names in the
coordinate data set.

```{r}
# identify unmatched countries in the map_data  
x1 <- unique(final$country)[is.na(charmatch(unique(final$country), 
                                      unique(coordinate$region), 
                                      nomatch = NA_integer_))]

x1
```

### 2. Find correspondant country names

There are currently 28 countries unmatched country names at the "final"
data set. We need to find the correspondent names in the map data -
"coordinate" data set.

```{r, warning=FALSE}
# find correpondent names for unmatched countries at the "final" data set 
newx1<- c("Barbuda", "Bolivia", "Brunei", "Ivory Coast", "Republic of Congo", 
          "Cape Verde", "Czech Republic", "Micronesia", "UK", "China", 
          "Iran", "Nevis", "South Korea", "Laos", "Moldova", "South Korea",  
          "Palestine", "Russia", "Swaziland", "Syria", "Trinidad", "Turkey", 
          "Tuvalu", "Tanzania", "USA", "Saint Vincent", "Venezuela", "Vietnam")

new_names <- data.frame(cbind(oldCountry=x1,
                              newCountry= newx1))
head(new_names,5)

```

### 3. Replace unmatched country names

We need to utilize the **stri_replace_all_fixed** function from the
{stringi} package to replace the mismatched country names.

```{r}
library(stringi)

final2 <- final %>% 
  mutate(newCountry = ifelse(country %in% new_names$oldCountry, 
                            new_names$newCountry[match(country, new_names$oldCountry)], 
                            country)) %>%   
  mutate(newCountry = as.character(newCountry))

```

\# Note: This code uses the **`ifelse()`** function to check if each
value in **`final$newCountry`** is in the **`new_names$oldCountry`**
vector. If it is, it replaces the value with the corresponding value in
**`new_names$newCountry`**. If it is not, it keeps the original value.

## **4. Merge map data with population data**

```{r, warning=FALSE}
# Use a variable for the year to make it easier to change the year in the future
Year <- 2020

library(tidyverse)
world <- final2 %>%
  merge(coordinate, by.x = "newCountry", by.y = "region") %>%
  filter(year == Year) %>%
  arrange(group, order) 


world_map <- ggplot(data = world) +
geom_map(
aes(map_id = region,
    x = long,
    y = lat,
    fill = PopGrowthRate),
    map = world) +
  labs(title = paste("Population growth rate in", Year),
       subtitle = paste(length(unique(world$newCountry)), "countries")) +
  ggthemes::theme_map() + 
  scale_fill_gradient2(name = "Poupation Growth Rate") 

world_map 
```

## The change of fertility rate

```{r, warning = FALSE, message=FALSE}

g <- final %>% 
   drop_na (hdicode) %>%  
  rename(HDI = hdicode) %>% 
  ggplot(aes(x = hdi,
             y = fertilityRate,
             color = HDI),
         alpha = 0.01,
         size = population) +
  geom_point() +
  geom_smooth(se = FALSE) +
  xlim(0, 1) +
  labs(title="Country fertility rate by human development index \n (1990-2021)",
       x = "Human Development Index",
       y = "Fertility Rate") 

print(g)
```

```{r, warning = FALSE, message = FALSE}
pacman::p_load(gganimate, gifski, png, av, tweenr)
# Creating an animating graph 
g + 
  transition_time(as.integer(year)) +
  labs(title = "Change of Fertility rate by Human Development Index (1990-2020) "
       ,subtitle = "Year: {frame_time}")
# + facet_wrap(~HDI)
```

The graph suggests that some countries have experienced a significant
decline in fertility rate. To further investigate this trend, I plan to
calculate the year-over-year change in fertility rate for each country.

```{r}
# calculate fertility rate difference
Diffence <- final %>%  
  group_by(iso3) %>%  
  arrange(iso3, year) %>%  
  mutate( Diff = fertilityRate - lag(fertilityRate) ) %>%  
  select(country, iso3, year, hdi, hdicode, fertilityRate, Diff, region) 

d <- Diffence %>%  
  na.omit() %>%  
  ggplot(aes(x = hdicode, y = Diff,color = hdicode)) +
  geom_violin() 

d

d + 
  transition_time(as.integer(year)) +
  labs(title =  "Fertility rate change by year",
    subtitle = "Year: {frame_time}")

```

I am interested in identifying which countries experienced the largest
decline in fertility rate for each year between 1990 and 2020, and how
this trend relates to the country's human development index ranking

```{r}
Diffence |> 
    filter(year == 2020) %>%
  group_by(year) %>%
  slice_min(Diff, n = 10) %>%
  drop_na (hdicode) %>% 
  arrange(Diff) %>%
  ggplot(aes(y = country, x = Diff,
             color = hdicode,
             size = abs(Diff))) +
  geom_point() +
  #  remove scale legend
  guides(size = "none") +
  labs(title = "Countries with biggest fertility rate drop ", 
       subtitle = paste("Year: 2020" ),
       x = "Fertility rate drop") 
```

# Case study

**identify countries to invest for pharmaceutical companies (by market
size and potential)**

To further utilize the data, let's consider a practical scenario where
pharmaceutical companies want to identify countries for investment
opportunities. To do this, we can use a model that considers key factors
such as market size and potential.

In terms of market size, we can analyze population size, life
expectancy, and fertility rate data from the United Nations, Department
of Economic and Social Affairs, Population Division (2022). World
Population Prospects 2022.

In terms of potential, we can analyze Gross National Income (GNI) per
capita data from the HDR21-22_Composite_indices_metadata.csv file, which
can be found on the UNDP website. We should also transpose the GNI per
capita data to make it more useful for our analysis.

## Metadata

1.  **life expectancy and annual population**

**File name**: "1950-2100, medium (ZIP, 7.77 MB)"

**Source link**:
[**https://population.un.org/wpp/Download/Standard/CSV/**](https://population.un.org/wpp/Download/Standard/CSV/)

-   Variables

    -   *LEx*: Life Expectancy at Birth, both sexes (years)

    -   *TPopulation1July*: Total Population, as of 1 July (thousands)

2.  **Gross national income (GNI) per capita**

    **File name**: HDR21-22_Composite_indices_metadata.csv

    **Source link**:
    <https://hdr.undp.org/data-center/documentation-and-downloads>

-   Variables

    -   *gnipc*: Gross national income (GNI) per capita

**Transpose the Gross national income (GNI) per capita data**

```{r}
library(dplyr)

GNI <- HDI_raw %>%  
  select(c(1:2,134:165)) |>
  pivot_longer(cols = starts_with("gnipc_"),
               names_to = "year",
               names_prefix = "gnipc_",
               values_to = "gni") %>%  
  na.omit()

project <- merge(final,GNI) 
```

### Identify countries with a large market size and potential

**Step 1: Population size**

We can first filter out those with lower life expectancy and higher
population growth rates. One strategy for this is to only consider
countries with population growth rates below the median of all
countries.

![]()

```{r, warning=FALSE}
countries <-  project  %>%  
  group_by(year, hdicode) %>%
  filter(hdicode != "Low" & 
           PopGrowthRate < median(PopGrowthRate) &
           population > median(population))

listCountry <- countries %>%
  pull(country) %>%
  unique()

df <-  world %>% 
  filter(country %in% listCountry)

ggplot(df) +
geom_map(aes(map_id = region,
    x = long,
    y = lat,
    fill = PopGrowthRate),
    map = df ) +
  labs(title = paste("Countries with low population growth rate"),
       subtitle = paste(length(unique(df$newCountry)), "countries")) +
  ggthemes::theme_map() + 
  scale_fill_gradient(name = "Poupation Growth Rate")
```

**Step 2: filter countries by the Gross national income (GNI) per capita
data**

```{r}
countries <- countries |> 
  group_by(year)  |> 
  filter(gni > mean(gni)) 

listCountry <- countries |> 
  pull(country) |> 
  unique()

listCountry

df <-  world %>% 
  filter(country %in% listCountry)

ggplot(df) +
geom_map(aes(map_id = region,
    x = long,
    y = lat,
    fill = PopGrowthRate),
    map = df ) +
  labs(title = paste("Countries with low population growth rate"),
       subtitle = paste(length(unique(df$newCountry)), "countries")) +
  ggthemes::theme_map() + 
  scale_fill_gradient(name = "Poupation Growth Rate")

```

**Step 3: check which countries have bigger population\
**

```{r}
if (!require("ggrepel")) install.packages("ggrepel")
library(ggrepel)

graph_project <- countries %>% 
  filter(year %in% c(2010:2020)) %>%
  ggplot(aes(PopGrowthRate, year, color=iso3)) +
  scale_y_continuous(breaks = c(2010:2021)) + 
  geom_point(aes(size = population,na.rm = TRUE)) +
  guides(size = FALSE, color = FALSE) +
  labs(x = "Population Growth Rate", y = "Year")

# add country which have population bigger than the median on 2020 
Year <- 2020

graph_project +
  geom_label_repel(data = countries %>% 
                     filter(year == Year & 
                              population > median(population)),
                   aes(label = country),
                  nudge_x = 0.5,
                  label.size = 0.01,
                  max.overlaps = getOption("ggrepel.max.overlaps", 
                                           default = 100),
                  na.rm = TRUE) +
  labs(title = "Countries with population bigger than median",
       subtitle = paste("Year : ", Year))


```

# Conclusion

This project identifies 24 countries that might need health care
services or medical cares for the elderly people considering the bigger
life expectancy, low population growth rate and greater gross national
income per capital.

```{r}
countries %>%
  pull(country) %>%
  unique()
```

# Limitations

Despite having a list of potential countries in need of medical services
for the elderly, this project does not accurately identify the specific
countries to invest in based on life expectancy, population growth rate,
and gross national income per capita. To enhance the applicability of
the conclusion, I need to develop or utilize a model for more precise
analysis.

## Reference:

-   UN Human Development Index:
    <https://hdr.undp.org/data-center/human-development-index#/indicies/HDI>

-   Details on how the HDI is calculated
    <http://hdr.undp.org/sites/default/files/hdr2022_technical_notes.pdf>

-   Gapminder foundation:
    <https://www.gapminder.org/tools/#$chart-type=bubbles&url=v1>
